---
title: "Forecasting categorical time series"
description: "Visualizing time-tracking data from 2022 to 2024"
draft: true
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.showtext=TRUE, eval = FALSE)
```

```{r libraries}
library(stargazer)
library(tidyverse)
library(readr)
library(patchwork)
library(ggtext)
library(showtext)
library(emojifont)
library(grid)
library(transformr)
library(lubridate)
library(plotly)
library(scales)
library(ggiraph)
library(glue)
library(knitr)
```

```{r theme}
font_add_google("Fira Sans", "firasans")
font_add(family = "Font Awesome 6 Brands",
                   regular = "data/Font Awesome 6 Brands-Regular-400.otf")

# showtext::showtext_auto()


# blog_theme <-
#    theme_minimal(18, base_family = "firasans") +
#    theme(
#      plot.background = element_rect(fill = "white", color = NA),
#      plot.title.position = "plot",
#      plot.title = element_text(size = 24, margin = margin(b = 1, unit = "line")),
#      legend.position = c(0, 1),
#      legend.direction = "horizontal",
#      legend.justification = c(0, 1),
#      legend.title.align = 1,
#      axis.title.y = element_text(hjust = 0, size = 10),
#      axis.title.x = element_text(hjust = 0, size = 10),
#      panel.grid.major = element_line(color = "#d3d9db"),
#      panel.grid.minor = element_blank()
#    )
#  
# theme_set(blog_theme)
```

```{r}
data_long20 <- read.csv("data/data_2020.csv") %>%
  mutate(
    datetime = as.POSIXct(datetime, format="%m/%d/%Y %H:%M", tz="GMT+3"),
    day_of_week = weekdays(datetime),
    time_column = if_else(substr(as.character(datetime), 12, 16) == "", "00:00", substr(as.character(datetime), 12, 16)),
    hour = hour(datetime),
    day = wday(datetime, label = TRUE, abbr = TRUE, week_start = 1),
    month = month(datetime, label = TRUE, abbr = FALSE),
    activity = case_when(
     tolower(activity) %in% c("misc", "mic") ~ "Miscellaneous",
     activity %in% c("Transport", "Transportation") ~ "Going Somewhere",
     activity == "Communication" ~ "Social Life",
     activity == "TM" ~ "Toastmasters",
     TRUE ~ activity))


data <- read_delim("data/KSE_life_2.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(time = col_time(format = "%H:%M")), 
    trim_ws = TRUE)

data_long <- gather(data, day, activity, -time) %>% 
  na.omit() %>%
  mutate(day = ifelse(day == "9/30/2022", "10/1/2022", day))

data_long$datetime <- as.POSIXct(paste(data_long$day, data_long$time), format="%m/%d/%Y %H:%M:%S", tz="GMT+3")

data_long <- data_long %>%
   mutate(
     date = as.Date(datetime),
     day_of_month = mday(datetime),
   time_column = if_else(substr(as.character(datetime), 12, 16) == "", "00:00", substr(as.character(datetime), 12, 16)),
   activity = case_when(
     tolower(activity) %in% c("misc", "mic") ~ "Miscellaneous",
     activity %in% c("Transport", "Transportation") ~ "Going Somewhere",
     activity == "Communication" ~ "Social Life",
     activity == "TM" ~ "Toastmasters",
     TRUE ~ activity),
     hour = hour(datetime),
     #day = wday(datetime, label = TRUE, abbr = TRUE, week_start = 1),
     month = month(datetime, label = TRUE, abbr = FALSE)
   )

activity_colors <- c(
   "Sleep" = "#3B4CC0",  # Deep Blue (Night)
   "Social Life" = "#87CEEB",  # SkyBlue
   "Sport" = "#1D8E3A",  # Forest Green (Active)
   "Work" = "#C0392B",  # Red (Energy)
   "Reading" = "#F39C12",  # Golden Orange (Focus)
   "Japanese" = "#FF69B4",  # HotPink
   "Study" = "#FFD700",  # Gold
   "Food" = "#99CC99",  # Soft Green
   "Waste" = "#333333",  # Dark Charcoal
   "Piano" = "purple",  # HotPink
   "Planning" = "#FF6347",  # Tomato
   "Toastmasters" = "#8E44AD",  # Deep Purple (Creative)
   "Going Somewhere" = "#95A5A6",  # Urban Gray
   "Miscellaneous" = "#27AE60"  # Fresh Green
)

data1 <- data_long20 %>%
  select(datetime, activity) %>%
  mutate(index = row_number())

# Select relevant columns from the second dataset and add an index
data2 <- data_long %>%
  select(datetime, activity) %>%
  mutate(index = row_number())

# Combine the two datasets
combined_data <- bind_rows(data1, data2, .id = "source") %>%
  select(index, datetime, activity) %>% 
  mutate(
    period = case_when(
     year(datetime) %in% 2019:2020 ~ "2019-2020",
     year(datetime) %in% 2022:2024 ~ "2022-2024",
     TRUE ~ NA_character_ ),
    date = as.Date(datetime),
    time_column = if_else(substr(as.character(datetime), 12, 16) == "", "00:00", substr(as.character(datetime), 12, 16))) %>% 
  select(-index)

```

## Problem

## Data

```{r}
combined_data %>%
  filter(period == "2022-2024") %>%
  ggplot(aes(date, time_column, fill = activity)) +
  geom_tile() +
  scale_fill_manual(values = activity_colors) +
  scale_x_date() +
   labs(title = NULL, x = NULL, y = NULL) +
   theme_minimal() +
   theme(
     legend.position = "none",
     panel.grid = element_blank(),
     plot.title = element_blank(),
     axis.text.x = element_text(angle = 0, size = 10, face = "bold"),
     axis.text.y = element_text(hjust = 0.5, size = 5, face = "bold")
     )
```

We can think about different activities as signals. So Sleep could be a signal that is high during the night and low during the day. Social Life could be a signal that is high during the evening and low during the day. Work could be a signal that is high during the day and low during the night.

```{r}
ts_data <- combined_data %>%
  mutate(ts = case_when(
   activity == "Sleep" ~ 0, 
   activity == "Miscellaneous" ~ 1,
    activity == "Social Life" ~ 2, 
    activity == "Study" ~ 3,
    activity == "Work" ~ 4,
    activity == "Going Somewhere" ~ 5, 
    activity == "Toastmasters" ~ 6,
    activity == "Sport" ~ 7, 
    activity == "Reading" ~ 8, 
    activity == "Japanese" ~ 9,
    activity == "Food" ~ 10,
    activity == "Waste" ~ 11,
    activity == "Piano" ~ 12, 
    activity == "Planning" ~ 13,
    TRUE ~ NA))
```

```{r}
ggplot(ts_data %>% 
  mutate(month = month(datetime),
         year = year(datetime)) %>% 
  filter(year == 2023, month == 1, date == "2023-01-01"), 
       aes(datetime, ts)) +
  geom_line() +
  labs(title = NULL, x = NULL, y = NULL) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.title = element_blank(),
    axis.text.x = element_text(angle = 0, size = 10, face = "bold"),
    axis.text.y = element_text(hjust = 0.5, size = 5, face = "bold")
  )

  
```

Looks messy! Looks like you've made a mistake. Let's try to assign the

## How similar are days?

One way of approaching this is to treat each day as a vector. In our case that is a 72-item vector (24 hours \* 3 activities). We can then calculate the cosine similarity between two days.

Cosine Similarity

Got the idea from here: https://www.relevantmisc.com/python/munging/2020/06/23/discretizing-timeseries/

## What was I doing in 2021-2022?

```{r}
ts_complete <- ts_data %>% 
  complete(datetime = seq(min(datetime), max(datetime), by = "20 mins"), fill = list(activity_code = NA)) 

ts_reg <- ts_data %>% 
  mutate(year = year(datetime),
         month = month(datetime, label = TRUE, abbr = FALSE),
         wday = wday(datetime, label = TRUE, abbr = FALSE),
         mday = mday(datetime),
         week = week(datetime),
         hour = hour(datetime),
         minute = minute(datetime),
         lag_1 = lag(ts, 1),
         lag_2 = lag(ts, 2),
         lag_3 = lag(ts, 3),
         lead_1 = lead(ts, 1),
         lead_2 = lead(ts, 2),
         lead_3 = lead(ts, 3)) %>% 
  filter(year %in% 2022:2024) %>%
  arrange(datetime)


glimpse(ts_reg)
```

```{r}
# Fit a logistic regression model
logistic_model <- glm(activity_binary ~ wday + month + hour + lag_1 + lag_2 + lag_3 + lead_1 + lead_2 + lead_3,
                      data = ts_reg %>% 
                        mutate(activity_binary = ifelse(ts_reg$activity == "Sleep", 1, 0)),
                      family = binomial(link = "logit"))

stargazer::stargazer(logistic_model, type = "text")
```


```{r}
library(depmixS4)

# We’ll assume we have K hidden states. Choose K based on domain knowledge or model selection.
K <- 13 

# depmix uses a formula interface. For categorical data, we use "multinomial" family.
# The response is 'activity' and we assume it follows a multinomial distribution.
# Each hidden state has its own emission probabilities for each activity category.

# If activity is a factor with L levels, multinomial responses can be specified like this:
model <- depmixS4::depmix(ts ~ 1, data = ts_reg, nstates = K, family = multinomial())

# Fit the model
set.seed(123) # For reproducibility
hmm_fit <- fit(model)

# Summary of the fitted model
summary(hmm_fit)
```








```{r}
library(broom)
library(rstanarm)
library(bayestestR)
library(bayesplot)
library(insight)

mod <- lm(ts ~ wday*(lag_1 + lag_2 + lag_3 + lead_1 + lead_2 + lead_3), 
          data = ts_reg %>% 
            filter(ts == 10))

stargazer::stargazer(mod, type = "text")
```

```{r}
modb <- stan_glm(
  ts ~ year*(lag_1 + lag_2 + lag_3 + lead_1 + lead_2 + lead_3), 
  data = ts_reg %>% 
    filter(ts == ),
  seed=111
)
```

